# "Penguins"

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(palmerpenguins)
library(quarto)

```

A simple example based on Allison Horst's [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/) dataset. Here we look at how penguin body mass varies across both sex and species (use the provided inputs to filter the dataset by bill length and island):

```{ojs}
//| echo = false
data = FileAttachment("data/palmer-penguins.csv").csv({ typed: true })
```

```{ojs }
//| echo = false
viewof bill_length_mm = Inputs.range(
  [32, 50], 
  {value: 35, step: 1, label: "Bill length (min):"}
)
viewof islands = Inputs.checkbox(
  ["Torgersen", "Biscoe", "Dream"], 
  { value: ["Torgersen", "Biscoe"], 
    label: "Islands:"
  }
)
```

```{ojs}
//| echo: false
filtered = data.filter(function(penguin) {
  return bill_length_mm < penguin.bill_length_mm &&
         islands.includes(penguin.island);
})
```

<!-- ::: {.panel-tabset} -->

## Plot

```{ojs}
//| echo: false
//| output: true
//| label: fig-penguin-body-mass
//| fig-cap: "Penguin body mass by sex and species"
Plot.rectY(filtered, 
  Plot.binX(
    {y: "count"}, 
    {x: "body_mass_g", fill: "species", thresholds: 20}
  ))
  .plot({
    facet: {
      data: filtered,
      x: "sex",
      y: "species",
      marginRight: 80
    },
    marks: [
      Plot.frame(),
    ]
  }
)
```

## Data

The `penguins` data from the [**palmerpenguins**](https://allisonhorst.github.io/palmerpenguins "palmerpenguins R package") package contains size measurements for `r nrow(penguins)` penguins from three species observed on three islands in the Palmer Archipelago, Antarctica./ The plot below shows the relationship between flipper and bill lengths of these penguins.

```{r}
#| label: plot-penguins
#| warning: false
#| echo: false

ggplot(penguins, 
       aes(x = flipper_length_mm, y = bill_length_mm)) +
  geom_point(aes(color = species, shape = species)) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(
    title = "Flipper and bill length",
    subtitle = "Dimensions for penguins at Palmer Station LTER",
    x = "Flipper length (mm)", y = "Bill length (mm)",
    color = "Penguin species", shape = "Penguin species"
  ) +
  theme_minimal()
```
